---
title: "Chapter 3 Introduction to Hypothesis Testing"
author: "Chihara-Hesterberg"
date: "November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
library(ggplot2)
library(dplyr)
```

##Permutation Tests

###Section 3.3
```{r}
Beerwings <- read.csv("https://sites.google.com/site/chiharahesterberg/data2/Beerwings.csv")

Beerwings %>% group_by(Gender) %>% summarize(mean(Hotwings))

observed <- 14.5333- 9.3333 #store observed mean differences

#Get hotwings variable
hotwings <- pull(Beerwings, Hotwings)

#alternatively, using "piping" operator
hotwings <- Beerwings %>% pull(Hotwings)
```


```{r}
#set.seed(0)
 N <- 10^5 - 1  #set number of times to repeat this process
 result <- numeric(N) # space to save the random differences
 for(i in 1:N)
{
  index <- sample(30, size = 15, replace = FALSE) # sample of numbers from 1:30
  result[i] <- mean(hotwings[index]) - mean(hotwings[-index])
}

ggplot() + geom_histogram(aes(result)) + geom_vline(xintercept=observed, colour = "red")

```

Another visualization of distribution.
```{r}
ggplot() + stat_ecdf(aes(result)) + geom_vline(xintercept = observed, colour = "red")
```

Compute P-value.
```{r}
(sum(result >= observed) + 1)/(N + 1)  #P-value
```


###Example 3.4 Verizon
###Permutation test

```{r}
Verizon <- read.csv("https://sites.google.com/site/chiharahesterberg/data2/Verizon.csv")

Verizon %>% group_by(Group) %>% summarize(mean(Time))

Time <- Verizon %>% pull(Time)
Time.ILEC <- Verizon %>% filter(Group == "ILEC") %>% pull(Time)
Time.CLEC <- Verizon %>% filter(Group == "CLEC") %>% pull(Time)

observed <- mean(Time.ILEC) - mean(Time.CLEC)
observed

N <- 10^4-1  #set number of times to repeat this process
#set.seed(99)
result <- numeric(N) # space to save the random differences
for(i in 1:N)
{
  index <- sample(1687, size = 1664, replace = FALSE) #sample of numbers from 1:1687
  result[i] <- mean(Time[index]) - mean(Time[-index])
}

ggplot() + geom_histogram(aes(result)) + geom_vline(xintercept=observed, colour="red")+
 xlab("xbar1-xbar2")

(sum(result <= observed) + 1)/(N + 1)  #P-value

```

###Example 3.6, Verizon cont.

median, trimmed means

```{r}
Verizon %>% group_by(Group) %>% summarize(median(Time))

#Difference in means
observed <- median(Time.ILEC) - median(Time.CLEC)
observed

#Differnce in trimmed means
observed2 <- mean(Time.ILEC, trim = .25) - mean(Time.CLEC, trim = .25)
observed2

N <- 10^4-1  #set number of times to repeat this process
#set.seed(99)
result  <- numeric(N) # space to save the random differences
result2 <- numeric(N)
for(i in 1:N)
{
  index <- sample(1687, size=1664, replace = FALSE) #sample of numbers from 1:1687
  result[i] <- median(Time[index]) - median(Time[-index])
  result2[i] <- mean(Time[index], trim = .25) - mean(Time[-index], trim = .25)
}

ggplot() + geom_histogram(aes(result)) + geom_vline(xintercept=observed, colour="red") +
   xlab("median1-median2")

#P-value difference in means
(sum(result <= observed) + 1)/(N+ 1)

ggplot() + geom_histogram(aes(result2)) + geom_vline(xintercept=observed2, colour="red") +   xlab("trimmedMean1-trimmedMean2")

#P-value difference in trimmed means
(sum(result2 <= observed2) + 1)/(N+ 1)
```

###Example 3.6, Verzion continued

difference in proportion of time > 10
and ratio of variances
```{r}
observed3 <- mean(Time.ILEC > 10) - mean(Time.CLEC > 10)
observed3

#ratio of variances
observed4 <- var(Time.ILEC)/var(Time.CLEC)
observed4

N <- 10^4-1  #set number of times to repeat this process
#set.seed(99)
 result3 <- numeric(N)
 result4 <- numeric(N)

 for(i in 1:N) {
  index <- sample(1687, size = 1664, replace = FALSE)
  result3[i] <- mean(Time[index] > 10) - mean(Time[-index] > 10)
  result4[i] <- var(Time[index])/var(Time[-index])
  }

 ggplot() + geom_histogram(aes(result3)) + geom_vline(xintercept=observed3, colour="red") + xlab("difference in proportions")

#P-value difference in means
(sum(result3 <= observed3) + 1)/(N+ 1)
 
```

Results for ratio of variances:
```{r}
ggplot() + geom_histogram(aes(result4)) + geom_vline(xintercept=observed4, colour="red") +   xlab("variance1/variance2")

(sum(result4 <= observed4) + 1)/(N+ 1)  #P-value
```

###Example 3.8
```{r}
Recidivism <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Recidivism.csv")
```
There are some missing values in the variable <tt> `Age25`</tt>. The <tt> `complete.cases`</tt> command gives the row numbers where values are not missing. We create a new data frame omitting the rows where there is a missing value in the <tt> 'Age25' </tt> variable.
```{r}
temp <- Recidivism %>% filter(complete.cases(Age25))
```

 Also, the variable <tt>`Recid`</tt> is a factor variable coded "Yes" or "No". We convert it to a numeric variable of 1's and 0's. The command below results in a new column added to <tt>temp</tt> called <tt>'Recid2`</tt>.
```{r}
temp <- temp %>% mutate(Recid2 = ifelse(Recid=="Yes", 1, 0))
head(temp, 2)   #inspect first few rows

temp %>% group_by(Age25) %>% summarize(mean(Recid2))

observed <- .365 - .306

Recid2 <- pull(temp, Recid2)   #extract column vector 
N <- 10^4 - 1
result <- numeric(N)

for (i in 1:N)
{
  index <- sample(17019, size = 3077, replace = FALSE)
  result[i] <- mean(Recid2[index]) - mean(Recid2[-index])
}

2* (sum(result >= observed) + 1)/(N + 1)
```


###Section 3.4 Matched Pairs

Results from a diving championship.

```{r}
Diving2017 <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Diving2017.csv")

Diving2017 <- Diving2017 %>% mutate(Diff = Final - Semifinal)
head(Diving2017)   #inspect modified data frame

Diff <- pull(Diving2017, Diff)  #extract vector

#Diff <- Diving %>% pull(Diff)   #alternative syntax

observed <- mean(Diff)

N <- 10^5 - 1
result <- numeric(N)

for (i in 1:N)
{
  Sign <- sample(c(-1, 1), 12, replace = TRUE)
  Diff2 <- Sign*Diff
  result[i] <- mean(Diff2)
}

ggplot() + geom_histogram(aes(result)) + geom_vline(xintercept = observed, colour="red") + xlab("Difference")

2* (sum(result >= observed) + 1)/(N + 1)
```

