---
title: "Chapter 4 Sampling Distributions"
author: "Chihara-Hesterberg"
date:  "November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
library(dplyr)
library(ggplot2)
```

###Example 4.2: Sampling distribution from Exp(1/15)
```{r}
Xbar <- numeric(1000)
#set.seed(300)
for (i in 1:1000)
{
  x <- rexp(100, rate = 1/15)
  Xbar[i] <- mean(x)
}

ggplot() + geom_histogram(aes(Xbar), bins = 15) + xlab("means")

df <- data.frame(Xbar)
ggplot(df, aes(sample=Xbar)) + stat_qq() + stat_qq_line()

mean(Xbar)
sd(Xbar)
```

###Example 4.3: Sampling Dist from Unif[0,1]
```{r}
maxY <- numeric(1000)
#set.seed(100)
for (i in 1:1000)
 {
   y <- runif(12)        #draw random sample of size 12
   maxY[i] <- max(y)     #find max, save in position i
 }

ggplot() + geom_histogram(aes(maxY), binwidth=.05, center=.975) + xlab("maximums") 

```

To create a histogram with a density curve imposed, we will need to create a data frame that holds the <tt>'maxY'</tt> variable. We also create a function for the density curve $f(x)=12x^{11}$.


```{r}
df <- data.frame(maxY)
myfun <- function(x){12*x^{11}}

ggplot(df) + geom_histogram(aes(maxY, y = stat(density)), binwidth=.05, center=.975) +xlab("maximums") + stat_function(fun = myfun)
```

###Example 4.6 Sum of Poisson random variables
```{r}
X <- rpois(10^4, 5)   #Draw 10^4 values from Pois(5)
Y <- rpois(10^4, 12)   #Draw 10^4 values from Pois(12)
W <- X + Y

df1 <- data.frame(W)
df2 <- data.frame(x=2:35, y = dpois(2:35,17))
ggplot(df1, aes(W)) + geom_histogram(aes(y=stat(density)), bins=12) + geom_line(data=df2, aes(x=x, y=y), colour = "red")

mean(W)
var(W)
```

###Example 4.7
Sampling distribution simulation
Sample of size 30 from gamma r=5, lambda=2

```{r}
#set.seed(10)
Xbar <- numeric(1000)
for (i in 1:1000)
  {
    x <- rgamma(30, shape = 5, rate = 2)
    Xbar[i] <- mean(x)
  }

ggplot() + geom_histogram(aes(Xbar), bins=15) + labs(title = "Distribution of means")

ggplot() + stat_qq(aes(sample = Xbar)) 

#If you want a line, then
df <- data.frame(Xbar)
ggplot(df, aes(sample = Xbar)) + stat_qq() + stat_qq_line()

mean(Xbar)
sd(Xbar)
sum(Xbar > 3)/1000
#alternatively
mean(Xbar > 3)
```

###Example 4.11 R Note

```{r}
dbinom(25, 120, .3)

pbinom(25, 120, .3)
```