---
title: "Chapter 5 Bootstrap"
author: "Chihara-Hesterberg"
date: "November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "50%")
library(dplyr)
library(ggplot2)
```

###Example 5.2
```{r}
my.sample <- rgamma(16, 1, 1/2)

N <- 10^5
my.boot <- numeric(N)
for (i in 1:N)
 {
  x <- sample(my.sample, 16, replace = TRUE)  #draw resample
  my.boot[i] <- mean(x)                     #compute mean, store in my.boot
  }

ggplot() + geom_histogram(aes(my.boot), bins=15)

mean(my.boot)  #mean
sd(my.boot)    #bootstrap SE
```

###Example 5.3

Arsenic in wells in Bangladesh
```{r}
Bangladesh <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Bangladesh.csv")

ggplot(Bangladesh, aes(Arsenic)) + geom_histogram(bins = 15)

ggplot(Bangladesh, aes(sample = Arsenic)) + stat_qq() + stat_qq_line()

Arsenic <- pull(Bangladesh, Arsenic)
#Alternatively
#Arsenic <- Bangladesh$Arsenic

n <- length(Arsenic)
N <- 10^4

arsenic.mean <- numeric(N)

for (i in 1:N)
{
   x <- sample(Arsenic, n, replace = TRUE)
   arsenic.mean[i] <- mean(x)
}

ggplot() + geom_histogram(aes(arsenic.mean), bins = 15) + 
  labs(title="Bootstrap distribution of means") + 
  geom_vline(xintercept = mean(Arsenic), colour = "blue")

df <- data.frame(x = arsenic.mean)
ggplot(df, aes(sample = x)) + stat_qq() + stat_qq_line()

mean(arsenic.mean)                 #bootstrap mean
mean(arsenic.mean) - mean(Arsenic) #bias
sd(arsenic.mean)                   #bootstrap SE

sum(arsenic.mean > 161.3224)/N
sum(arsenic.mean < 89.75262)/N
```

###Example 5.4 Skateboard
```{r}
Skateboard <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Skateboard.csv")

testF <- Skateboard %>% filter(Experimenter == "Female") %>% pull(Testosterone)
testM <- Skateboard %>% filter(Experimenter == "Male") %>% pull(Testosterone)

observed <- mean(testF) - mean(testM)

nf <- length(testF)
nm <- length(testM)

N <- 10^4

TestMean <- numeric(N)

for (i in 1:N)
{
  sampleF <- sample(testF, nf, replace = TRUE)
  sampleM <- sample(testM, nm, replace = TRUE)
  TestMean[i] <- mean(sampleF) - mean(sampleM)
}

df <- data.frame(TestMean)
ggplot(df) + geom_histogram(aes(TestMean), bins = 15) + 
  labs(title = "Bootstrap distribution of difference in means", xlab = "means") +
  geom_vline(xintercept = observed, colour = "blue")

ggplot(df, aes(sample = TestMean))  + stat_qq() + stat_qq_line()

mean(testF) - mean(testM)
mean(TestMean)
sd(TestMean)

quantile(TestMean,c(0.025,0.975))

mean(TestMean)- observed  #bias
```

###Permutation test for Skateboard means
```{r}
testAll <- pull(Skateboard, Testosterone)
#testAll <- Skateboard$Testosterone

N <- 10^4 - 1  #set number of times to repeat this process

#set.seed(99)
result <- numeric(N) # space to save the random differences
for(i in 1:N)
  {
  index <- sample(71, size = nf, replace = FALSE) #sample of numbers from 1:71
  result[i] <- mean(testAll[index]) - mean(testAll[-index])
}

(sum(result >= observed)+1)/(N + 1)  #P-value

ggplot() + geom_histogram(aes(result), bins = 15) + 
  labs(x = "xbar1-xbar2", title="Permutation distribution for testosterone levels") +
  geom_vline(xintercept = observed, colour = "blue")

df <- data.frame(result)
ggplot(df, aes(sample = result)) + stat_qq() + stat_qq_line()

```

###Section 5.4.1 
Matched pairs for Diving data

```{r}
Diving2017 <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Diving2017.csv")
Diff <- Diving2017 %>% mutate(Diff = Final - Semifinal) %>% pull(Diff)
#alternatively
#Diff <- Diving2017$Final - Diving2017$Semifinal
n <- length(Diff)

N <- 10^5
result <- numeric(N)

for (i in 1:N)
{
  dive.sample <- sample(Diff, n, replace = TRUE)
  result[i] <- mean(dive.sample)
}

ggplot() + geom_histogram(aes(result), bins = 15)

quantile(result, c(0.025, 0.975))
```


###Example 5.5 
Verizon cont.
Bootstrap means for the ILEC data and for the CLEC data

Bootstrap difference of means.
```{r}
Verizon <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Verizon.csv")

Time.ILEC <- Verizon %>% filter(Group == "ILEC") %>% pull(Time)
Time.CLEC <- Verizon %>% filter(Group == "CLEC") %>% pull(Time)

observed <- mean(Time.ILEC) - mean(Time.CLEC)

n.ILEC <- length(Time.ILEC)
n.CLEC <- length(Time.CLEC)

N <- 10^4

time.ILEC.boot <- numeric(N)
time.CLEC.boot <- numeric(N)
time.diff.mean <- numeric(N)

set.seed(100)
for (i in 1:N)
 {
  ILEC.sample <- sample(Time.ILEC, n.ILEC, replace = TRUE)
  CLEC.sample <- sample(Time.CLEC, n.CLEC, replace = TRUE)
  time.ILEC.boot[i] <- mean(ILEC.sample)
  time.CLEC.boot[i] <- mean(CLEC.sample)
  time.diff.mean[i] <- mean(ILEC.sample) - mean(CLEC.sample)
}

#bootstrap for ILEC
ggplot() + geom_histogram(aes(time.ILEC.boot), bins = 15) + 
  labs(title = "Bootstrap distribution of ILEC means", x = "means") + 
  geom_vline(xintercept = mean(Time.ILEC), colour = "blue") + 
  geom_vline(xintercept = mean(time.ILEC.boot), colour = "red", lty=2)

summary(time.ILEC.boot)

df <- data.frame(x = time.ILEC.boot)
ggplot(df, aes(sample = x)) + stat_qq() + stat_qq_line()

#bootstrap for CLEC
ggplot() + geom_histogram(aes(time.CLEC.boot), bins = 15) + 
  labs(title = "Bootstrap distribution of CLEC means", x = "means") + 
  geom_vline(xintercept = mean(Time.CLEC), colour = "blue") + 
  geom_vline(xintercept = mean(time.CLEC.boot), colour = "red", lty = 2)

df <- data.frame(x = time.CLEC.boot)
ggplot(df, aes(sample = x)) + stat_qq() + stat_qq_line()

#Different in means
ggplot() + geom_histogram(aes(time.diff.mean), bins = 15) + 
  labs(title = "Bootstrap distribution of difference in means", x = "means") +
  geom_vline(xintercept = mean(time.diff.mean), colour = "blue") + 
  geom_vline(xintercept = mean(observed), colour = "red", lty = 2)

df <- data.frame(x = time.diff.mean)
ggplot(df, aes(sample = x)) + stat_qq() + stat_qq_line()

mean(time.diff.mean)
quantile(time.diff.mean, c(0.025, 0.975))

```

###Section 5.5 
Verizon cont.

Bootstrap difference in trimmed means

```{r}
Time.ILEC <- Verizon %>% filter(Group == "ILEC") %>% pull(Time)
Time.CLEC <- Verizon %>% filter(Group == "CLEC") %>% pull(Time)
n.ILEC <- length(Time.ILEC)
n.CLEC <- length(Time.CLEC)

N <- 10^4
time.diff.trim <- numeric(N)

#set.seed(100)
for (i in 1:N)
{
  x.ILEC <- sample(Time.ILEC, n.ILEC, replace = TRUE)
  x.CLEC <- sample(Time.CLEC, n.CLEC, replace = TRUE)
  time.diff.trim[i] <- mean(x.ILEC, trim = .25) - mean(x.CLEC, trim = .25)
}

ggplot() + geom_histogram(aes(time.diff.trim), bins = 15) + 
  labs(x = "difference in trimmed means") + 
  geom_vline(xintercept = mean(time.diff.trim),colour = "blue") + 
  geom_vline(xintercept = mean(Time.ILEC, trim = .25) - mean(Time.CLEC, trim = .25), colour = "red", lty = 2)

df <- data.frame(x = time.diff.trim)
ggplot(df, aes(sample = x)) + stat_qq() + stat_qq_line()

mean(time.diff.trim)
quantile(time.diff.trim, c(0.025,0.975))
```

###Section 5.5 Other statistics
Verizon cont: 

Bootstrap of the ratio of means

<tt>`Time.ILEC`</tt> and <tt>`Time.CLEC`</tt> created above.

<tt>`n.ILEC`</tt>, <tt>`n.CLEC`</tt> created above

```{r}
N <- 10^4
time.ratio.mean <- numeric(N)

#set.seed(100)
for (i in 1:N)
{
  ILEC.sample <- sample(Time.ILEC, n.ILEC, replace = TRUE)
  CLEC.sample <- sample(Time.CLEC, n.CLEC, replace = TRUE)
  time.ratio.mean[i] <- mean(ILEC.sample)/mean(CLEC.sample)
}

ggplot() + geom_histogram(aes(time.ratio.mean), bins = 12) + 
  labs(title = "bootstrap distribution of ratio of means", x = "ratio of means") +
  geom_vline(xintercept = mean(time.ratio.mean), colour = "red", lty = 2) + 
  geom_vline(xintercept  = mean(Time.ILEC)/mean(Time.CLEC), col = "blue")

df <- data.frame(x = time.ratio.mean)
ggplot(df, aes(sample = x)) + stat_qq() + stat_qq_line()

mean(time.ratio.mean)
sd(time.ratio.mean)
quantile(time.ratio.mean, c(0.025, 0.975))
```

###Example 5.7 Relative risk example
```{r}
highbp <- rep(c(1,0),c(55,3283))   #high blood pressure
lowbp <- rep(c(1,0),c(21,2655))    #low blood pressure

N <- 10^4
boot.rr <- numeric(N)
high.prop <- numeric(N)
low.prop <- numeric(N)

for (i in 1:N)
{
   x.high <- sample(highbp,3338, replace = TRUE)
   x.low  <- sample(lowbp, 2676, replace = TRUE)
   high.prop[i] <- sum(x.high)/3338
   low.prop[i]  <- sum(x.low)/2676
   boot.rr[i] <- high.prop[i]/low.prop[i]
}

ci <- quantile(boot.rr, c(0.025, 0.975))

ggplot() + geom_histogram(aes(boot.rr), bins = 15) + 
  labs(title = "Bootstrap distribution of relative risk", x = "relative risk") +
  geom_vline(aes(xintercept = mean(boot.rr), colour = "mean of bootstrap")) +
  geom_vline(aes(xintercept = 2.12, colour="observed rr"), lty = 2) + 
  scale_colour_manual(name="", values = c("mean of bootstrap"="blue", "observed rr" = "red"))

temp <- ifelse(high.prop < 1.31775*low.prop, 1, 0)
temp2 <- ifelse(high.prop > 3.687*low.prop, 1, 0)
temp3 <- temp + temp2

df <- data.frame(y=high.prop, x=low.prop, temp, temp2, temp3)
df1 <- df %>% filter(temp == 1)
df2 <- df %>% filter (temp2 == 1)
df3 <- df %>% filter(temp3 == 0)

ggplot(df, aes(x=x, y = y)) + 
  geom_point(data =df1, aes(x= x, y = y), colour = "green") + 
  geom_point(data = df2, aes(x = x, y = y), colour = "green") + 
  geom_point(data = df3, aes(x = x, y = y), colour = "red") + 
  geom_vline(aes(xintercept = mean(low.prop)), colour = "red") +
  geom_hline(yintercept = mean(high.prop), colour = "red") + 
  geom_abline(aes(intercept = 0, slope = 2.12, colour = "observed rr"), lty = 2, lwd = 1) + 
  geom_abline(aes(intercept = 0, slope = ci[1], colour = "bootstrap CI"), lty = 2, lwd = 1) + 
  geom_abline(intercept = 0, slope = ci[2], colour = "blue", lty = 2, lwd = 1) +
  scale_colour_manual(name="", values=c("observed rr"="black", "bootstrap CI" = "blue")) +
  labs(x = "Proportion in low blood pressure group", y = "Proportion in high blood pressure group")
```