---
title: "Chapter 3 Introduction to Hypothesis Testing"
author: "Chihara-Hesterberg"
date: "July 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
```

##Permutation Tests

###Section 3.3
```{r}
Beerwings <- read.csv("https://sites.google.com/site/chiharahesterberg/data2/Beerwings.csv")

tapply(Beerwings$Hotwings, Beerwings$Gender, mean)

observed <- 14.5333- 9.3333 #store observed mean differences

#Get hotwings variable
hotwings <- Beerwings$Hotwings
```

An alternative way to extract the vector:
```{r}
hotwings <- subset(Beerwings, select = Hotwings, drop = TRUE)
```
The <tt>`drop = TRUE`</tt> argument converts the hotwings to a vector. Without this, hotwings would be a 30x1 data frame.

```{r}
#set.seed(0)
 N <- 10^5 - 1  #set number of times to repeat this process
 result <- numeric(N) # space to save the random differences
 for(i in 1:N)
{
  index <- sample(30, size = 15, replace = FALSE) # sample of numbers from 1:30
  result[i] <- mean(hotwings[index]) - mean(hotwings[-index])
}


hist(result, xlab = "xbarM - xbarF", main = "Permutation distribution for hot wings")
abline(v = observed, col = "blue", lty = 5)
```

Another visualization of distribution.
```{r}
plot.ecdf(result)
abline(v = observed, col = "blue", lty = 5)
```

Compute P-value.
```{r}
(sum(result >= observed) + 1)/(N + 1)  #P-value
```


###Example 3.4 Verizon
###Permutation test

```{r}
Verizon <- read.csv("https://sites.google.com/site/chiharahesterberg/data2/Verizon.csv")

tapply(Verizon$Time, Verizon$Group, mean)


Time <- subset(Verizon, select = Time, drop = T)
Time.ILEC <- subset(Verizon, select = Time, Group == "ILEC", drop = TRUE)
Time.CLEC <- subset(Verizon, select = Time, Group == "CLEC", drop = TRUE)

observed <- mean(Time.ILEC) - mean(Time.CLEC)
observed


N <- 10^4-1  #set number of times to repeat this process
#set.seed(99)
result <- numeric(N) # space to save the random differences
for(i in 1:N)
{
  index <- sample(1687, size = 1664, replace = FALSE) #sample of numbers from 1:1687
  result[i] <- mean(Time[index]) - mean(Time[-index])
}

hist(result, xlab = "xbar1 - xbar2",
      main = "Permutation Distribution for Verizon repair times")
abline(v = observed, col = "blue", lty = 5)

(sum(result <= observed) + 1)/(N + 1)  #P-value

```

###Example 3.6, Verizon cont.

median, trimmed means

```{r}
tapply(Verizon$Time, Verizon$Group, median)

#Difference in means
observed <- median(Time.ILEC) - median(Time.CLEC)
observed

#Differnce in trimmed means
observed2 <- mean(Time.ILEC, trim = .25) - mean(Time.CLEC, trim = .25)
observed2

N <- 10^4-1  #set number of times to repeat this process
#set.seed(99)
result  <- numeric(N) # space to save the random differences
result2 <- numeric(N)
for(i in 1:N)
{
  index <- sample(1687, size=1664, replace = FALSE) #sample of numbers from 1:1687
  result[i] <- median(Time[index]) - median(Time[-index])
  result2[i] <- mean(Time[index], trim = .25) - mean(Time[-index], trim = .25)
}

hist(result, xlab = "median1 - median2",
    main = "Permutation Distribution for medians")
abline(v = observed, col = "blue", lty = 5)

#P-value difference in means
(sum(result <= observed) + 1)/(N+ 1)


hist(result2, xlab = "trimMean1 - trimMean2",
     main = "Permutation Distribution for trimmed means")
abline(v = observed, col = "blue", lty = 5)

#P-value difference in trimmed means
(sum(result2 <= observed2) + 1)/(N+ 1)
```

###Example 3.6, Verzion continued

difference in proportion of time > 10
and ratio of variances
```{r}
observed3 <- mean(Time.ILEC > 10) - mean(Time.CLEC > 10)
observed3

#ratio of variances
observed4 <- var(Time.ILEC)/var(Time.CLEC)
observed4

N <- 10^4-1  #set number of times to repeat this process
#set.seed(99)
 result3 <- numeric(N)
 result4 <- numeric(N)

 for(i in 1:N) {
  index <- sample(1687, size = 1664, replace = FALSE)
  result3[i] <- mean(Time[index] > 10) - mean(Time[-index] > 10)
  result4[i] <- var(Time[index])/var(Time[-index])
  }

hist(result3, xlab = "Difference in proportions",  main = "Repair times > 10 hours")
abline(v = observed3, lty = 5, col = "blue")

(sum(result3 <= observed3) + 1)/(N+ 1)  #P-value
```

Results for ratio of variances:
```{r}
hist(result4, xlab = "variance1/variance2",  main = "Ratio of variances")
abline(v = observed4, lty = 5, col = "blue")

(sum(result4 <= observed4) + 1)/(N+ 1)  #P-value
```

###Example 3.8
```{r}
Recidivism <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Recidivism.csv")
```
There are some missing values in the variable <tt> `Age25`</tt>. The <tt> `complete.cases`</tt> command gives the row numbers where values are not missing.
```{r}
k <- complete.cases(Recidivism$Age25)
```

 Also, the variable <tt>`Recid`</tt> is a factor variable coded "Yes" or "No". We convert it to a numeric variable of 1's and 0's.
```{r}
Recid2 <- ifelse(Recidivism$Recid[k] == "Yes", 1, 0)
Age25.2 <- Recidivism$Age25[k]

table(Age25.2)
tapply(Recid2, Age25.2, mean)

observed <- .365 - .306

N <- 10^4 - 1
result <- numeric(N)

for (i in 1:N)
{
  index <- sample(17019, size = 3077, replace = FALSE)
  result[i] <- mean(Recid2[index]) - mean(Recid2[-index])
}

2* (sum(result >= observed) + 1)/(N + 1)
```


###Section 3.4 Matched Pairs

Results from a diving championship.

```{r}
Diving2017 <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Diving2017.csv")

Diff <- Diving2017$Final - Diving2017$Semifinal
observed <- mean(Diff)

N <- 10^5 - 1
result <- numeric(N)

for (i in 1:N)
{
  Sign <- sample(c(-1, 1), 12, replace = TRUE)
  Diff2 <- Sign*Diff
  result[i] <- mean(Diff2)
}

hist(result)
abline(v = observed, col = "blue")

2* (sum(result >= observed) + 1)/(N + 1)
```

