---
title: "Chap 10 Categorical Data"
author: "Chihara-Hesterberg"
date: "December 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "50%")
library(ggplot2)
library(dplyr)
```

###Section 10.2

Here is a function that computes the chi-square test statistic

The code below gives a function that is a bit more enhanced than the code in the textbook:
```{r}
chisq <- function(observed, print = TRUE) {
  # Chi-square statistic for independence in a contingency table,
  # with related data exploration.
  # observed is the observed contingency table

  observedWithTotals <- cbind(observed, total = rowSums(observed))
  observedWithTotals <- rbind(observedWithTotals, total = colSums(observedWithTotals))
  expected <- outer(rowSums(observed), colSums(observed)) / sum(observed)
  statistic <- sum((observed-expected)^2/expected)
  if (print){
  cat("Observed, with totals:\n")
  print(observedWithTotals)
  cat("\nRow Fractions:\n")
  print(round(observed / rowSums(observed), 3))
  cat("\nColumn Fractions:\n")
  print(round(observed / rep(colSums(observed), each = nrow(observed)), 3))

  cat("\nExpected:\n")
  print(round(expected, 1))
  cat("\nDifference:\n")
  print(round(observed - expected, 1))

  cat("\nChi-squared statistic for independence:", round(statistic, 1), "\n")
}
  return(invisible(statistic))
}
```
Import the General Social Survey data. We are interested in the two variables, <tt>`Education`</tt> and <tt>`DeathPenalty`</tt>.

Using the <tt>`str()`</tt> command, we note that these two variables have missing values. We will create a new data frame which contains just the two variables of interest and also, only contains the rows without NA's.

```{r}
GSS2002 <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/GSS2002.csv")

str(GSS2002)

df <- GSS2002 %>% 
       select(Education, DeathPenalty) %>% 
       filter(!is.na(Education) & !is.na(DeathPenalty))

Education <- pull(df, Education)
DeathPenalty <- pull(df, DeathPenalty)

table(Education, DeathPenalty)
```

<tt>`Education`</tt> is a factor variable and the default ordering of the levels is alphabetical. We use the <tt>`ordered`</tt> command to *order* the levels.

```{r}
Education <- ordered(Education, levels = c("Left HS", "HS", "Jr Col", "Bachelors", "Graduate"))                    
table(Education, DeathPenalty)
```


Use function created above to calculate chi-square test statistic

```{r}
observedChi2 <- chisq(table(Education, DeathPenalty))
observedChi2
```

Now run the permutation test

```{r}
N <- 10^4-1
result<-numeric(N)

for (i in 1:N)
 {
   DP.permutation <-sample(DeathPenalty)
   GSS.table <- table(Education, DP.permutation)
   result[i]<-chisq(GSS.table, print = FALSE)
 }

ggplot() + geom_histogram(aes(result)) + 
   labs(title = "Distribution of chi-square statistics", x = "chi-square statistic") + 
   geom_vline(xintercept = observedChi2, colour = "blue")

(sum(result >= observedChi2) + 1)/(N + 1)
```


Optional: Create a histogram with the density curve
imposed onto the histogram. The <tt>ggplot()</tt> command will require a data frame which contains the variable of interest.

```{r}
df <- data.frame(result)
ggplot(df) + geom_histogram(aes(result, y = stat(density))) + 
   labs(title = "Distribution of chi-square statistics", x = "chi-square statistic") +   
   geom_vline(xintercept = observedChi2, colour = "blue") +
   stat_function(fun = dchisq, args = list(df = 4), colour = "green")
```

The <tt>`chisq.test`</tt> command also has an option that will perform this permutation test:

```{r}
chisq.test(Education, DeathPenalty, simulate.p.value = TRUE, B = 10^4 - 1)
mat <- table(Education, DeathPenalty)
chisq.test(mat, simulate.p.value = TRUE, B = 10^4-1)
```

###Example 10.2
```{r}
mat <- rbind(c(42, 50), c(30, 87))
chisq.test(mat)
```

###Section 10.3.3 Fisher's Exact Test

```{r}
fisher.test(mat)
```

###Section 10.4 Test of Homogeneity
```{r}
candy.mat <- rbind(c(42, 20, 38), c(33, 27, 50))
candy.mat

chisq.test(candy.mat)
```

###Section 10.6
```{r}
Phillies2009 <- read.csv("http://sites.google.com/site/chiharahesterberg/data2/Phillies2009.csv")
Homeruns <- pull(Phillies2009, Homeruns)

lambda <- mean(Homeruns)
dpois(0:5, lambda)
table(Homeruns)

table(Homeruns)/162
```