---
title: "Introduction to R, part 2"
author: "Chihara-Hesterberg"
date: "July 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
library(resampledata)
library(ggplot2)
library(dplyr)
```


### Vectors in R

The basic data object in R is the vector.
Even scalars are vectors of length 1.

There are several ways to create vectors.

The <tt>:</tt> operator creates sequences incrementing/decrementing
by 1.

```{r}
1:10
5:-3
```

The <tt>seq</tt> function creates sequences also.
```{r}
seq(0, 3, by = .2)
seq(0, 3, length = 15)
```

To create vectors with no particular pattern, use the
<tt>c()</tt> function (<tt>c</tt> for **c**ombine).

```{r}
c(1, 4, 8, 2, 9)
x <- c(2, 0, -4)
x
c(x, 0:5, x)
```

For vectors of characters,

```{r}
c("a", "b", "c", "d")
```

or logical values (note that there are no double quotes):

```{r}
c(TRUE, FALSE, FALSE, TRUE, TRUE, FALSE)
```

The <tt>rep</tt> command for repeating values:

```{r}
rep("a", 5)
rep(c("a", "b"), 5)
rep(c("a", "b"), c(5, 2))
```

### The  <tt>class</tt> attribute

Use <tt>data.class</tt> to determine the class attribute of an object.

```{r}
state.name
data.class(state.name)
state.name == "Idaho"
data.class(state.name == "Idaho")

head(FlightDelays$Carrier)
data.class(FlightDelays$Carrier)
```
  

### Basic Arithmetic

```{r}
x <- 1:5
x - 3
x*10
x/10
x^2
2^x
log(x)

w <- 6:10
w
x*w                 #coordinate-wise multiplication
```

#### Logical expressions

```{r}
x < 3
```

### Subsetting a vector

In many cases, we will want only a portion of a data set. For
subsetting a vector, the basic syntax is <tt>vector[*index*]</tt>.
In particular, note the use of *brackets* to indicate that we are
subsetting.

```{r}
state.name               # 50 states (alphabetical order)
state.name[c(1, 25, 50)] # the 1st, 25th, and 50th 
state.name[-(1:10)]      # remove the first 10.

z <- c(8, 3, 0, 9, 9, 2, 1, 3)
z
z[4]              # The fourth element of z
z[c(1, 3, 4)]     # The first, third and fourth element,
z[-c(1, 3, 4)]    # All elements except the first, third and fourth
```

To return the values of <tt>z</tt> less than 4, we first introduce the
<tt>which</tt> command:

```{r}
which(z < 4)             # which positions are z values < 4?
index <- which(z < 4)    # store in index
index
z[index]                 # return z[c(2, 3, 6, 7)]
```

Suppose you want to find those observations when the delay length
was greater than the mean delay length. We'll store this in a vector
called <tt>index</tt>.

```{r}
delay <- FlightDelays$Delay
index <- which(delay > mean(delay))
head(index)
```

Thus, observations in rows  2, 10, 12, 14, 15, 16 are the first
six that correspond to flights that had delays that were larger
than the average delay length.

### Extracting parts of a data frame

To subset particular rows of a data frame, use the <tt>filter</tt> command in the *dplyr* package.

For example, to create a data frame with just the United Airlines flights:
```{r}
United <- FlightDelays %>% filter(Carrier == "UA")
```
The <tt>select</tt> command in the **dplyr** package allows you to  extract just certain variables (columns). For example, to create a data frame containing just the <tt>Carrier</tt> and <tt>Delay</tt> variables:

```{r}
FlightDelays2 <- FlightDelays %>% select(Carrier, Delay)
```
Finally, we can combine these two actions to extract just certain rows and certain columns:

```{r}
United2 <- FlightDelays %>% filter(Carrier == "UA") %>% select(Carrier, Delay)
```

Now, suppose you want to work with a single variable in a data frame.

```{r}
delay <- FlightDelays %>% select(Delay)
head(delay)
mean(delay)
data.class(delay)
```
The problem is that in the above, the <tt>select</tt> command returns a data frame, and the <tt>mean</tt> command operates on vectors.

If we just want to extract one variable from a data frame and we want that variable to be a vector, use the <tt>pull</tt> command.

```{r}
delay <- FlightDelays %>% pull(Delay)
mean(delay)

#Alternatively, we have seen that the $ operator can be used
delay <- FlightDelays$Delay
```



