---
title: "Introduction to R, Exploratory Data Analysis"
author: "Chihara-Hesterberg"
date: "July 2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width="50%")
```

Type commands after the prompt $>$ and
then press the <tt>$<$ENTER$>$</tt> key.

Note: anything after the pound symbol # is a
comment---explanatory text that is not executed.

```{r}
4*9           # Simple arithmetic
```

Create a sequence incrementing by $1$:

```{r}
20:30
```

We will create an object called <tt>dog</tt> and assign it the values
1, 2, 3, 4, 5. The symbol $<$- is the assignment operator.

```{r}
dog <- 1:5
dog

dog + 10

3*dog

sum(dog)        # 1+2+3+4+5
```

The object <tt>dog</tt> is called a  *vector*.

If you need to abort a command, press the escape key  **<ESC>**.
The up arrow key $\uparrow$ can be used to recall previous entries.

To obtain help on any of the commands,  type the
name of the command you wish help on:

```{r}
?sum
```

### Importing data

Data for the third edition of the textbook can be downloaded from the web site [linked phrase] (http://github.com/chihara/MathStatsResamplingR/Edition3/Data) or by installing the R package **resampledata3**.

For instance, let's start with the Flight Delays Case Study
(see
Chapter 1, **Case Studies**), of the text for a description of this
data set).  We use the <tt>read.csv</tt> command to import the data into our R workspace:

```{r}
library(resampledata3)  

#Alternatively, if the data set has been downloaded to the working directory
#FlightDelays<- read.csv("FlightDelays.csv")
#View(FlightDelays)
```

The <tt>str</tt> command gives a compact display of the internal structure of the data frame:

```{r}
str(FlightDelays)
```
Alternatively, the <tt>summary</tt> gives numeric summaries of the variables:

```{r}
summary(FlightDelays)
```

Other commands to get basic information about the data frame include <tt>names</tt>, to view the names of the variables in <tt>FlightDelays</tt>:

```{r}
names(FlightDelays)
```

To view the first 6 rows of the data, use the <tt>head</tt> command:
```{r}
head(FlightDelays)     #default, 6 rows

head(FlightDelays, 4)  #first 10 rows
```

(What do you suppose the <tt>tail</tt> command does?)

The columns are the *variables*. There are two types of variables:  *numeric*,
for example, <tt>FlightLength</tt> and <tt>Delay</tt> and *factorI (also called *categorical*) (for example <tt>Carrier</tt> and <tt>DepartTime</tt>). The rows are called *observations* or *cases*.

To check the  size (number of rows and columns) of the data frame, type
```{r}
dim(FlightDelays)   # dim = dimension
```

This tells us that there are 4029 observations and 10 columns.

### Tables, bar charts and histograms

The factor variable <tt>Carrier</tt> in the <tt>FlightDelays</tt> data set assigns
each flight to one of two  *levels*: UA or AA. To obtain the summary of this variable, we
use the $ operator to extract this variable from the <tt>FlightDelays</tt> data frame:

```{r}
table(FlightDelays$Carrier)
```

#### Remark
R is **case-sensitive**! <tt>Carrier</tt> and <tt>carrier</tt> are considered different.

To create a contingency table summarizing the relationship between
carrier (<tt>Carrier</tt>) and whether or not a flight was delayed by more than 30 minutes (<tt>Delayed30</tt>), type:

```{r}
table(FlightDelays$Carrier, FlightDelays$Delayed30)
```

To perform additional operations, we will first store the table object:

```{r}
out <- table(FlightDelays$Carrier, FlightDelays$Delayed30)
out

addmargins(out)  # table with row/column sums 
```

The <tt>proportions</tt> command gives joint or conditional distributions:

```{r}
proportions(out)           #joint distribution (sum of all cells = 1)

proportions(out, 1)        #conditional distributions (row sums = 1)
```

Thus, 9.8% of flights were American Airlines and delayed by more than 30
minutes, whereas of all American Airline flights, 13.5% were
delayed by more than 30 minutes

To visualize the distribution of a factor variable, create a *bar chart*:

```{r}
barplot(table(FlightDelays$Carrier))
```

The <tt>barplot</tt> command is part of base R, that is, it comes with the default installation of R. Researchers have developed *packages* that enhance many of the basic R commands. One such package is the **ggplot2** package.

```{r}
library(ggplot2)

ggplot(FlightDelays, aes(x = Carrier)) + geom_bar()
```

Note that the syntax is different from the base R command for a bar plot: in particular, we do not first have to create a table.

To visualize the distribution of a numeric variable, create a histogram.

```{r}
hist(FlightDelays$Delay)    # base R

ggplot(FlightDelays, aes(x = Delay)) + geom_histogram()   # ggplot2
```

The shape of the distribution of this variable is *right-skewed*.

The default number of bins for the **ggplot2** histogram is 30. We'll change this here:

```{r}
ggplot(FlightDelays, aes(x = Delay)) + geom_histogram(bins = 8, color = "white")
```


### Numeric Summaries

Because it is a bit cumbersome to use the syntax <tt>FlightDelays$Delay</tt>
each time we want to work with the <tt>Delay</tt> variable, we will extract this variable
and store it in the (vector) object <tt>delay</tt>:

```{r}
delay <- FlightDelays$Delay
head(delay)
mean(delay)
mean(delay, trim = .05)   #trimmed mean
median(delay)
```

Other basic statistics:

```{r}
max(delay)
min(delay)
range(delay)
var(delay)           #variance
sd(delay)            #standard deviation
quantile(delay)      #quartiles
```

If you need the population variance (that is, denominator of 1/n instead of
1/(n-1)),

```{r}
n <- length(delay)
(n-1)/n*var(delay)
```

### Statistics with grouping by a categorical variable

To find statistics for a numeric variable grouped by a categorical variable, one option is to use the <tt>tapply</tt> command

```{r}
tapply(FlightDelays$Delay, FlightDelays$Carrier, mean)
tapply(FlightDelays$Delay, FlightDelays$Carrier, sd)
```
Thus, the mean length of an American Airlines flight delay was 10.1 minutes compared to 15.98 minutes for United Airlines. The standard deviations for AA and UA were 40.08 m and 45.1 m, respectively.

Another approach is to use the commands in the **dplyr** package

```{r}
library(dplyr)
FlightDelays %>% group_by(Carrier) %>% summarize(mean(Delay), sd(Delay))
```

The %>% is the piping operator in the **dplyr** package: it takes the data set <tt>FlightDelays</tt> and
“pipes” it to the command group by. The <tt>group_by</tt> command groups the observations by <tt>Carrier</tt>,
that is by AA or UA. Then we pipe this information to the summarize command and compute the
mean and standard deviation of the <tt>Delay</tt> variable for each of AA and UA flights.

### Boxplots

Boxplots give a visualization of the 5-number summary
of a variable.

```{r}
summary(delay)         #numeric summary
boxplot(delay)
```

To compare the distribution of a numeric variable across the levels of
a categorical (factor) variable:

```{r}
boxplot(Delay ~ Day, data = FlightDelays)           # base R

ggplot(FlightDelays, aes(x = Day, y = Delay)) + geom_boxplot()  # ggplot2
```

### Normal quantile plots

We create normal quantile plots to see if it is plausible that data come from a normal distribution. The <tt>ggplot</tt> command requires a data frame for it's first argument.
```{r}
x <- c(17.7, 22.6, 26.1, 28.3, 30, 31.2, 31.5, 33.5, 34.7, 36)
df <- data.frame(x = x)
ggplot(df, aes(sample = x)) + geom_qq() + geom_qq_line() +
  labs(x = "Theoretical quantiles", y = "Sample quantiles")

ggplot(NCBirths2004, aes(sample = Weight))+ geom_qq() +
  geom_qq_line()

ggplot(NCBirths2004, aes(sample = Weight)) + geom_qq() +
  geom_qq_line() + facet_grid( ~ Smoker)
```

Let's generate a random sample of size 100 from the normal distribution N(30, 6) and
create the normal-quantile plot. 

```{r}
my.sample <- rnorm(100, 30, 6)
df <- data.frame(my.sample)     #create data frame
ggplot(df, aes(sample = my.sample)) + geom_qq() + geom_qq_line()
```

### Empirical Cumulative Distribution Function (ECDF)

```{r}
x <- c(3, 6, 15, 15, 17, 19, 24)
df <- data.frame(x = x)
ggplot(df, aes(x = x)) + stat_ecdf()

df <- data.frame(x = rnorm(25))     # random sample from N(0,1)
ggplot(df, aes(x = x)) + stat_ecdf() +
  stat_function(fun = pnorm, color = "blue") +
  labs(x="", y = "F(x)")
```

Using the data from the Beer and Hotwings case study:
```{r}
ggplot(Beerwings, aes(x = Beer, linetype = Gender, color = Gender)) +
  stat_ecdf() +  labs(x = "Beer (oz)", y = "F(x)") +
  geom_vline(xintercept = 25, lty = 3) 
```

### Scatter plots

```{r}
ggplot(Beerwings, aes(x = Hotwings, y = Beer)) +
  geom_point()
```

We can also distinguish the two genders by adding the color
aesthetic:
```{r}
ggplot(Beerwings, aes(x = Hotwings, y = Beer,
       color = Gender)) + geom_point()
```

### Misc. Remarks

+ Functions in R are called by typing their name followed by arguments surrounded
by *parentheses*: ex. <tt>hist(delay)</tt>. Typing a function name without parentheses
will give the code for the function.

```{r}
sd
```

+ We saw earlier that we can assign names to data (we created a vector called <tt>dog</tt>.)
Names can be any length, must start with a letter, and may contain letters or numbers:

```{r}
fish25 <- 10:35
fish25
```
Certain names are **reserved** so be careful to not use them: <tt>cat</tt>, <tt>c</tt>,  <tt>t</tt>, <tt>T</tt>, <tt>F</tt>,...

To be safe, before making an assignment, type the name:
```{r, eval = FALSE}
whale
```
If you get the output <tt>Problem: Object "whale" not found</tt>, then it is safe to use <tt>whale</tt>!

```{r}
whale <- 200
objects()
rm(whale)
objects()
```

+ In general, R is space-insensitive.

```{r}
3 +4
3+   4
mean(3+5)
mean ( 3 + 5 )
```

BUT, the assignment operator must not have spaces!
<tt>$<$-</tt> is different from <tt>$<$ \, - </tt>

+ To quit, type <tt>q()</tt> at the prompt.

You will be given an option to save the workspace.

If you select  **Yes**:
all objects created in this session are saved to your working directory so
that the next time you start up <tt>R</tt>, if you load this working directory,
these objects will still be available.
You will not have to re-import <tt>FlightDelays</tt>, for instance, nor recreate
<tt>delay</tt>.

You can, for back-up purposes, save  data to an external file/disk by using,
for instance, the  <tt>write.csv</tt> command. See the help file for more information.

